## Solutions to Questions

Below you will find the solutions to the questions from the preclass and inclass activities for this chapter. Only look at them after giving the questions a good try and speaking to the tutor about any issues.

### InClass Activities

#### Task 2: Give your Report a Heading

`r hide("Solution Task 2")`
```{r, echo=FALSE, results='asis'}
cat("You should have used only one hashtag to give the biggest heading size.

`# Lab 1 - The magnitude of the Ponzo Illusion varies as a function of Age`

")
```
`r unhide()`
**click the tab to see the solution**
<br>

#### Task 3: Creating a Code Chunk

`r hide("Solution Task 3")`
```{r, echo=FALSE, results='asis'}

cat("The `echo` rule should be set to `false`, as should the `warning` rule and `message` rule.

The code chunk line should look like: {r load_data, echo = FALSE, warning = FALSE, message = FALSE}.

Note we are not showing the back-ticks here but you will need to include them.")
```
`r unhide()`
**Click the tab to see the solution**
<br>

#### Task 4: Writing your Report

`r hide("Solution Task 4")`
```{r, echo = FALSE, results='asis'}

cat("Task 4 is about setting a title to `Header 2` style. This is done via two `##` at the start of the line - before the word Introduction in this case but dont forget the space. Worth noting that in normal scripts of R a `#` would result in turning the line into a comment. Here it sets it as `Header 2` much like a Word document header

Set an ordered list by putting `1` followed by a `.` then a space. And so on. It will only work if there is a empty line above the list as such:

1. The Perspective Hypothesis
2. The Framing Hypothesis
")
```
`r unhide()`
**Click the tab to see the solution**
<br>

#### Task 5: Making Text Bold or Italicized

`r hide("Solution Task 5")`
```{r, echo = FALSE, results='asis'}
cat("
To turn text to bold you need to put two `**` at the start and end of the word or sentence you want as bold. 
")
```
`r unhide()`
**Click the tab to see the solution**
<br>

#### Task 6: Adding Links to the Data in your Methods

`r hide("Solution Task 6")`
```{r, echo = FALSE, results='asis'}
cat("
To set a header as `Header 2` style use `##` at the start of the line.
To set a header as `Header 3` style use `###` at the start of the line.

A link is created by putting the words you want to act as the link between `[]` and then the link immediately after in `()` e.g. 
`[Lebowitz and Judisch (2016)]( https://www.jstor.org/stable/1420548?seq=1#page_scan_tab_contents)`
")
```
`r unhide()`
**Click the tab to see the solution**
<br>

#### Task 7: Adding an Image to your Methods

`r hide("Solution Task 7")`
```{r, echo = FALSE, results='asis'}
cat("
To set a header as `Header 3` style use `###` at the start of the line.

An image is created by putting the words you want to act as the name of the image `[]` and then the link to the image immediately after in `()`. The key thing is to start with an exclamation mark `!` e.g. 
`![name](link)`
")
```
`r unhide()`
**Click the tab to see the solution**
<br>

#### Task 8: Adding a Table to your Results

`r hide("Solution Task 8")`
```{r, echo = FALSE, results='asis'}
cat("
To set a header as `Header 2` style use `##` at the start of the line.

code chunk heading should read as {r table, echo = FALSE}
")
```
`r unhide()`
**Click the tab to see the solution**
<br>

#### Task 9: Adding a Figure to your Results

`r hide("Solution Task 9")`
```{r, echo = FALSE, results='asis'}
cat("
code chunk heading should read as {r plot, include = TRUE}
")
```
`r unhide()`
**Click the tab to see the solution**
<br>


