## Solutions to Questions

```{r lab2-solutions-data, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
# preclass
pong_data <- read_csv("./data/02-s01/preclass/PongBlueRedBack 1-16 Codebook.csv")
# inclass
responses <- read_csv("./data/02-s01/inclass/responses.csv")
qformats <- read_csv("./data/02-s01/inclass/qformats.csv")
scoring <- read_csv("./data/02-s01/inclass/scoring.csv")
pinfo <- read_csv("./data/02-s01/inclass/pinfo.csv")
# assignment
screening <- read_csv("./data/02-s01/homework/p_screen.csv")
responses <- read_csv("./data/02-s01/homework/QHC_responses.csv")
scoring <- read_csv("./data/02-s01/homework/QHC_scoring.csv")
``` 

Below you will find the solutions to the questions for the Activities for this chapter. Only look at them after giving the questions a good try and speaking to the tutor about any issues.

### PreClass Activities

#### PreClass Task 1

```{r Ch2PreClassTask1, eval = FALSE}
# To include variables:
select(pong_data, Participant, PaddleLength, TrialNumber, BackgroundColor, HitOrMiss)

# To exclude variables:
select(pong_data, -JudgedSpeed, -BallSpeed, -BlockNumber)

# To select in order:
select(pong_data, BallSpeed, HitOrMiss, JudgedSpeed, Participant, TrialNumber)
```

[Return to Task](#Ch2PreClassQueT1)

#### PreClass Task 2

```{r Ch2PreClassTask2, eval = FALSE}
arrange(pong_data, desc(HitOrMiss), desc(JudgedSpeed))
```

[Return to Task](#Ch2PreClassQueT2)

#### PreClass Task 3

``` {r Ch2PreClassTask3, eval = FALSE}
filter(pong_data, 
       JudgedSpeed == 1, 
       BallSpeed %in% c("2", "4", "5", "7"), 
       HitOrMiss == 0)
```

[Return to Task](#Ch2PreClassQueT3)

#### PreClass Task 4

``` {r Ch2PreClassTask4, eval = FALSE}
pong_data_filt <- filter(pong_data, TrialNumber >= 2) 
# remember you can call this variable anything, as long as it makes sense to yourself and others

pong_data2 <- mutate(pong_data_filt, TrialNumber = TrialNumber - 1)

```

[Return to Task](#Ch2PreClassQueT4)

#### PreClass Task 5

``` {r Ch2PreClassTask5, eval = FALSE}
group_by(pong_data2, BlockNumber, BackgroundColor)
```

[Return to Task](#Ch2PreClassQueT5)

#### PreClass Task 6

``` {r Ch2PreClassTask6, eval = FALSE}
pong_data2_group <- group_by(pong_data2, BackgroundColor, PaddleLength)
pong_data2_hits <- summarise(pong_data2_group, total_hits = sum(HitOrMiss))
```

* You should find that the number of hits made with the small paddle (50) and the red color background is `r pong_data2 %>% group_by(BackgroundColor, PaddleLength) %>% summarise(total_hits = sum(HitOrMiss)) %>% filter(BackgroundColor == "red") %>% slice(1) %>% pull(total_hits)`

[Return to Task](#Ch2PreClassQueT6)

### InClass Actitivies

#### InClass Task 2

```{r Ch2InClassTask2, eval = FALSE}
library(tidyverse)
# or
library("tidyverse")

# both do the same thing.
```

[Return to Task](#Ch2InClassQueT2)

#### InClass Task 3

```{r Ch2InClassTask3, eval = FALSE}
responses <- read_csv("responses.csv")                  
qformats <- read_csv("qformats.csv")                 
scoring <- read_csv("scoring.csv")                  
pinfo <- read_csv("pinfo.csv")
```

[Return to Task](#Ch2InClassQueT3)

#### InClass Task 7 

```{r Ch2InClassTask7, eval = FALSE}
aq_scores <- rscores %>% 
             group_by(Id) %>% # group by the ID number in column Id
             summarise(AQ = sum(Score)) # sum column Score to obtain AQ scores.
```

[Return to Task](#Ch2InClassQueT7)


#### InClass Task 8

```{r Ch2InClassTask8, eval = FALSE}
aq_scores2 <- responses %>% gather(Question, Response, Q1:Q10) %>% 
  inner_join(qformats, "Question") %>%
  inner_join(scoring, c("QFormat", "Response")) %>%
             group_by(Id) %>% 
             summarise(AQ = sum(Score))
```

[Return to Task](#Ch2InClassQueT8)

### Assignment Solution

```{r lab2-assignment-solutions-data, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
# assignment
screening <- read_csv("./data/02-s01/homework/p_screen.csv")
responses <- read_csv("./data/02-s01/homework/QHC_responses.csv")
scoring <- read_csv("./data/02-s01/homework/QHC_scoring.csv")
``` 

#### Assignment Task 1 - Oldest Participant

* Whether you coded this answer or just read from the data, the Participant with ID Number 3 is the oldest.

```{r Ch2AssignmentTask1}
oldest_participant <- 3

# This could also be answered with code. We haven't quite shown you how yet but it would look like this:
oldest_participant_code <- arrange(screening, desc(AGE)) %>% 
  slice(1) %>% 
  pull(ID)
```

[Return to Task](#Ch2AssignQueT1)

#### Assignment Task 2 - Arranging D-SPAN

* `arrange()` is the main function here
* You also needed to use `desc()` to sort from high to low

```{r Ch2AssignmentTask2}
cogtest_sort <- arrange(screening, desc(DSPAN))
```

[Return to Task](#Ch2AssignQueT2)

#### Assignment Task 3 - Foreign Language Speakers

1. First group the screening data by FLANG using `group_by()`
2. Next, summarise, paying attention to use the variable names as instructed

```{r Ch2AssignmentTask3}
screen_groups <- group_by(screening, FLANG) 
descriptives <- summarise(screen_groups, n = n(), median_age = median(AGE))
```

[Return to Task](#Ch2AssignQueT3)

#### Assignment Task 4 - Creating Percentage MOCA scores

* `mutate()` is the function to add a new column to data
* Here we are mutating/adding on a column called `MOCA_Perc` which shows a participant's MOCA score divided by 30 and multiplied by 100.

```{r Ch2AssignmentTask4}
screening <- mutate(screening, MOCA_Perc = (MOCA / 30) * 100)
```

[Return to Task](#Ch2AssignQueT4)

#### Assignment Task 5 - Remove the MOCA column

* `select()` is the key function to keep and remove certain columns.
* Two options here; both would give the same dataframe.
* The first option shows how to deselect a column and keep everything else.
* The second option shows how to select all the columns you want.
* Remember that order is very important and you should select the columns in the order you want.

```{r Ch2AssignmentTask5}
# First Option
screening <- select(screening, -MOCA)
# Second Option
screening <- select(screening, ID, AGE, SEX, HANDEDNESS, EDUCATION, MUSICAL, FLANG, DSPAN, D2, MOCA_Perc)
```

[Return to Task](#Ch2AssignQueT5)

#### Assignment Task 6 - Gather the Responses together

* `gather()` is the function to use here.
* People take a while to understand this function but spend sometime looking at the below example and it will start to make some sense.
* The first argument is the data. In this case `responses`.
* The second argument is the name of a new column which will hold all the question numbers (e.g. Q1, Q2, Q3, etc). This could be called anything but here we are calling it `Question`.
* The third argument is the name of a new column which will hold the answers to the questions stored in `Question` - e.g. the actual answer for that participant to Q1 will appear in this column. Again this could be called anything but here we are calling it `RESPONSE`.
* The last argument is the name of the columns you want to gather. Here we are gathering all columns between the Q1 column and the Q15 column.

```{r Ch2AssignmentTask6}
responses_long <- gather(responses, Question, RESPONSE, Q1:Q15)
```

[Return to Task](#Ch2AssignQueT6)

#### Assignment Task 7 - Joining the data

* `inner_join()` will combine all common information in two sets of data by a common column or columns.
* Here we are joining the data in `responses_long` with the data in `scoring` by the common column `RESPONSE`.
* Keep in mind that `inner_join()` keeps only the rows that have data in both datasets. It will remove rows that only has data in one dataset.
* When joining two datasets, join by **ALL** common columns when there are more than one column in common.

```{r Ch2AssignmentTask7}
responses_points <- inner_join(responses_long, scoring, "RESPONSE")
```

[Return to Task](#Ch2AssignQueT7)

#### Assignment Task 8 - Working the Pipes

* This is the code we started with.  

```{r Ch2AssignmentTask8_pre, eval = FALSE}
participant_groups <- group_by(responses_points, ID)
participant_scores <- summarise(participant_groups, Total_QHC = sum(SCORE))
participant_screening <- inner_join(participant_scores, screening, "ID")
screening_groups_new <- group_by(participant_screening, MUSICAL)
musical_means <- summarise(screening_groups_new, mean_score = mean(Total_QHC))
```

* Below is how to transcribe the above series of functions into a pipeline.
* Remember, when using pipes, the output of the previous function is the input of the subsequent function

```{r Ch2AssignmentTask8}
musical_means <- group_by(responses_points, ID) %>%
                  summarise(Total_QHC = sum(SCORE)) %>%
                  inner_join(screening, "ID") %>%
                  group_by(MUSICAL) %>%
                  summarise(mean_score = mean(Total_QHC))
```

[Return to Task](#Ch2AssignQueT8)

#### Assignment Task 9 - Difference in Musical Means

* People who play music have a QHC score that is 1.53 units higher than people who don't play music.
* You can do this by looking in `musical_means`, reading the values, and doing some quick maths.
* A second option is through code. Code is always better as it can reduce error and is reproducible!

```{r Ch2AssignmentTask9}
# Option 1
QHC_diff <- 1.53

# Option 2
# You will soon learn the functions to do this by code but here is how you could do it.
QHC_diff_code <- spread(musical_means, MUSICAL, mean_score) %>% 
  mutate(diff = YES - NO) %>% 
  pull(diff) %>% 
  round(2)
```

[Return to Task](#Ch2AssignQueT9)

<span style="font-size: 22px; font-weight: bold; color: var(--purple);">Chapter Complete!</span>
