## Solutions to Questions

```{r lab2-solutions-data, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
# preclass
pong_data <- read_csv("./data/02-s01/preclass/PongBlueRedBack 1-16 Codebook.csv")
# inclass
responses <- read_csv("./data/02-s01/inclass/responses.csv")
qformats <- read_csv("./data/02-s01/inclass/qformats.csv")
scoring <- read_csv("./data/02-s01/inclass/scoring.csv")
pinfo <- read_csv("./data/02-s01/inclass/pinfo.csv")
# assignment
screening <- read_csv("./data/02-s01/homework/p_screen.csv")
responses <- read_csv("./data/02-s01/homework/QHC_responses.csv")
scoring <- read_csv("./data/02-s01/homework/QHC_scoring.csv")
``` 

Below you will find the solutions to the questions for the Activities for this chapter. Only look at them after giving the questions a good try and speaking to the tutor about any issues.

### PreClass Activities

#### Task 1

`r hide("Solution Task 1")`
```{r eval = FALSE}
# To include variables:
select(pong_data, Participant, PaddleLength, TrialNumber, BackgroundColor, HitOrMiss)

# To exclude variables:
select(pong_data, -JudgedSpeed, -BallSpeed, -BlockNumber)

# To select in order:
select(pong_data, BallSpeed, HitOrMiss, JudgedSpeed, Participant, TrialNumber)
```
`r unhide()`  

**click the tab to see the solution**
<br>

#### Task 2

`r hide("Solution Task 2")`
```{r eval = FALSE}
arrange(pong_data, desc(HitOrMiss), desc(JudgedSpeed))
```
`r unhide()`  

**click the tab to see the solution**
<br>

#### Task 3

`r hide("Solution Task 3")`
``` {r eval = FALSE}
filter(pong_data, 
       JudgedSpeed == 1, 
       BallSpeed %in% c("2", "4", "5", "7"), 
       HitOrMiss == 0)
```
`r unhide()`  

**click the tab to see the solution**
<br>

#### Task 4

`r hide("Solution Task 4")`
``` {r eval = FALSE}
pong_data_filt <- filter(pong_data, TrialNumber >= 2) 
# remember you can call this variable anything, as long as it makes sense to yourself and others

pong_data <- mutate(pong_data_filt, TrialNumber = TrialNumber - 1)

```
`r unhide()`  

**click the tab to see the solution**
<br>


#### Task 5

`r hide("Solution Task 5")`
``` {r eval = FALSE}
group_by(pong_data, BlockNumber, BackgroundColor)
```
`r unhide()` 

**click the tab to see the solution**
<br>

#### Task 6

`r hide("Solution Task 6")`
``` {r eval = FALSE}
pong_data <- read_csv("PongBlueRedBack 1-16 Codebook.csv")
pong_data_group <- group_by(pong_data, BackgroundColor, PaddleLength)
pong_data_hits <- summarise(pong_data_group, total_hits = sum(HitOrMiss))
# the answer should give 517
```

### InClass Actitivies


#### Solution to Task 2

`r hide("Solution Task 2")`
```{r, eval = FALSE}
library(tidyverse)
# or
library("tidyverse")

# both do the same thing.
```
`r unhide()`

**Click the tab to see the solution**

#### Solution to Task 3

`r hide("Solution Task 3")`
```{r, eval = FALSE}
responses <- read_csv("responses.csv")                  
qformats <- read_csv("qformats.csv")                 
scoring <- read_csv("scoring.csv")                  
pinfo <- read_csv("pinfo.csv")
```
`r unhide()`

**Click the tab to see the solution**

#### Solution to Task 7

`r hide("Solution Task 7")`
```{r, eval = FALSE}
aq_scores <- rscores %>% 
             group_by(Id) %>% # group by the ID number in column Id
             summarise(AQ = sum(Score)) # sum column Score to obtain AQ scores.
```
`r unhide()`

**Click the tab to see the solution**

#### Solution to Task 8

`r hide("Solution Task 8")`
```{r, eval = FALSE}
aq_scores2 <- responses %>% gather(Question, Response, Q1:Q10) %>% 
  inner_join(qformats, "Question") %>%
  inner_join(scoring, c("QFormat", "Response")) %>%
             group_by(Id) %>% 
             summarise(AQ = sum(Score))
```
`r unhide()`

**Click the tab to see the solution**

### Assignment Solution

```{r lab2-assignment-solutions-data, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
# assignment
screening <- read_csv("./data/02-s01/homework/p_screen.csv")
responses <- read_csv("./data/02-s01/homework/QHC_responses.csv")
scoring <- read_csv("./data/02-s01/homework/QHC_scoring.csv")
``` 

#### Task 1 - Oldest Participant

Replace the `NULL` in the `T1` code chunk with the Participant ID of the oldest participant. Store this single value in `oldest_participant` (e.g. `oldest_participant <- 999`.

```{r T1}
oldest_participant <- 3

# This could also be answered with code. We haven't quite shown you how yet but it would look like this
oldest_participant_code <- arrange(screening, desc(AGE)) %>% 
  slice(1) %>% 
  pull(ID)

# Either way the Participant with ID = 3 is the oldest.
```

#### Task 2 - Arranging D-SPAN

Replace the `NULL` in the `T2` code chunk with code that arranges participants' D-SPAN performance from highest to lowest using the appropriate one-table dplyr (i.e., Wickham) verb. Store the output in `cogtest_sort`. (e.g. `cogtest_sort <- verb(data, argument)`)

```{r T2}
# arrange was the function and you also required desc() to sort high to low
cogtest_sort <- arrange(screening, desc(DSPAN))
```

#### Task 3 - Foreign Language Speakers

Replace the `NULL` in each of the two lines of code chunk `T3`, so that `descriptives` has a column called `n` that shows the number of participants that speak a foreign language and number of participants that do not speak a foreign language, and another column called `median_age` that shows the median age for those two groups. If you have done this correctly, `descriptives` should have 3 columns and 2 rows of data, not including the header row.

```{r T3}
# first run the group_by based on FLANG
screen_groups <- group_by(screening, FLANG)
# then summarise, paying attention to use the variable names as instructed
descriptives <- summarise(screen_groups, n = n(), median_age = median(AGE))
```

#### Task 4 - Creating Percentage MOCA scores

Replace the `NULL` in the `T4` code chunk with code using one of the dplyr verbs to add a new column called `MOCA_Perc` to the dataframe `screening` In this new column should be the `MOCA` scores coverted to percentages. The maximum achievable score on MOCA is `30` and percentages are calculated as `(participant score / max score) * 100`. Store this output in `screening`.

```{r T4}
# mutate is the necessary function to add a new column in a dataframe
screening <- mutate(screening, MOCA_Perc = (MOCA / 30) * 100)
```

#### Task 5 - Remove the MOCA column

Now that we have our MoCA score expressed as a percentage `MOCA_Perc` we no longer need the raw scores held in `MOCA`. Replace the `NULL` in the `T5` code chunk using a one-table dplyr verb to keep all the columns of `screening`, with the same order, but without the `MOCA` column. Store this output in `screening`.

```{r T5}
# Two options here; both would give the same dataframe
# All but MOCA
screening <- select(screening, -MOCA)
# All of these
screening <- select(screening, ID, AGE, SEX, HANDEDNESS, EDUCATION, MUSICAL, FLANG, DSPAN, D2, MOCA_Perc)
```

**Halfway There!**

The remaining tasks focus on merging two tables. 

You suspect that the older adults with musical experience might report more finely-tuned hearing abilities than those without musical experience. You therefore decide to check whether this trend exists in your data. You measured participant's self reported hearing abilties using the Better Hearing Institute Quick Hearing Check Questionnaire. In this questionnaire participants rated the extent to which they agree or disagree with a list of statements (e.g. 'I have a problem hearing over the telephone') using a 5 point Likert scale (Strongly Disagree, Slightly Disagree, Neutral, Slightly Agree, Strongly Agree). 

Each participant's response to each question is contained in the `responses` dataframe in your environment. Each response type is worth a certain number of points (e.g. Strongly Disagree = 0, Strongly Agree = 5), and the scoring key is contained in the `scoring` dataframe. A score for each participant is calculated by totalling up the number of points across all the questions to derive at an overall score. The lower the overall score, the better the participant's self-reported hearing ability.  

In order to score the questionnaire we first need to perform a couple of steps.

#### Task 6 - Gather the Responses together

Replace the `NULL` in the `T6` code chunk using code to gather the responses to all the questions of the QHC from wide format to tidy/long format. Name the first column `Question` and the second column `RESPONSE`. Store this output in `responses_long`.  

```{r T6}
# gather function as clued in the question; main thing is to state the columns
# and to name the columns accordingly
responses_long <- gather(responses, Question, RESPONSE, Q1:Q15)
```

#### Task 7 - Joining the data

Now we need to join the number of points for each response in `scoring` to the participants' responses in `responses_long`. 

Replace the `NULL` in the `T7` code chunk using `inner_join()` to combine `responses_long` and `scoring` into a new variable called `responses_points`. 

```{r T7}
# inner_join uses a common column/variable; in this case RESPONSE
responses_points <- inner_join(responses_long, scoring, "RESPONSE")
```

#### Task 8 - Working the Pipes

Below we have given you a code chunk with 5 lines of code. The code takes the data in its current long format and then creates a QHC score for each participant, before calculating a mean QHC score for the two groups of participants - those that play musical intruments and those that don't - and stores it in a variable called `musical_means`.  

```{r, eval = FALSE}
participant_groups <- group_by(responses_points, ID)
participant_scores <- summarise(participant_groups, Total_QHC = sum(SCORE))
participant_screening <- inner_join(participant_scores, screening, "ID")
screening_groups_new <- group_by(participant_screening, MUSICAL)
musical_means <- summarise(screening_groups_new, mean_score = mean(Total_QHC))
```

Replace the `NULL` in the `T8` code chunk with the following code converted into a fuctioning pipeline using pipes. Put each function on a new line one under the other. This pipeline should result in the mean QHC values of musical and non-musical people stored in `musical_means` which should be made of two rows by two columns.  

```{r T8}
# in pipes, the output of the previous function is the input of the subsequent function
musical_means <- group_by(responses_points, ID) %>%
                  summarise(Total_QHC = sum(SCORE)) %>%
                  inner_join(screening, "ID") %>%
                  group_by(MUSICAL) %>%
                  summarise(mean_score = mean(Total_QHC))
```

#### Task 9 - Difference in Musical Means

Finally, replace the `NULL` in the `T9` code chunk with a single value, to two decimal places, that is the value of how much higher the QHC score of people who play music is compared to people who don't play music (e.g. 2.93)

```{r T9}

QHC_diff <- 1.53

# Again this could more reproducibly be done by code but with functions you will soon learn. Here is how you would do it.

QHC_diff_code <- spread(musical_means, MUSICAL, mean_score) %>% 
  mutate(diff = YES - NO) %>% 
  pull(diff) %>% 
  round(2)

```

**End of Lab 2**
