## Solutions to Questions

```{r lab2-solutions-data, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
# preclass
pong_data <- read_csv("./data/02-s01/preclass/PongBlueRedBack 1-16 Codebook.csv")
# inclass
responses <- read_csv("./data/02-s01/inclass/responses.csv")
qformats <- read_csv("./data/02-s01/inclass/qformats.csv")
scoring <- read_csv("./data/02-s01/inclass/scoring.csv")
pinfo <- read_csv("./data/02-s01/inclass/pinfo.csv")
``` 

Below you will find the solutions to the questions for the Activities for this chapter. Only look at them after giving the questions a good try and speaking to the tutor about any issues.

### PreClass Activities

#### PreClass Task 1

* Using `select()` to include only stated columns:

```{r Ch2PreClassTask1-1, eval = FALSE}
select(pong_data, Participant, PaddleLength, TrialNumber, BackgroundColor, HitOrMiss)
```

* Using `select()` to exclude certain columns:

```{r Ch2PreClassTask1-2, eval = FALSE}
select(pong_data, -JudgedSpeed, -BallSpeed, -BlockNumber)
```

* Using `select()` to change the order of columns:

```{r Ch2PreClassTask1-3, eval = FALSE}
select(pong_data, BallSpeed, HitOrMiss, JudgedSpeed, Participant, TrialNumber)
```

[Return to Task](#Ch2PreClassQueT1)

#### PreClass Task 2

```{r Ch2PreClassTask2, eval = FALSE}
arrange(pong_data, desc(HitOrMiss), desc(JudgedSpeed))
```

[Return to Task](#Ch2PreClassQueT2)

#### PreClass Task 3

``` {r Ch2PreClassTask3, eval = FALSE}
filter(pong_data, 
       JudgedSpeed == 1, 
       BallSpeed %in% c("2", "4", "5", "7"), 
       HitOrMiss == 0)
```

[Return to Task](#Ch2PreClassQueT3)

#### PreClass Task 4

* first step is created with `filter()`
* the second step is created with `mutate()`

``` {r Ch2PreClassTask4, eval = FALSE}
pong_data_filt <- filter(pong_data, TrialNumber >= 2) 
pong_data2 <- mutate(pong_data_filt, TrialNumber = TrialNumber - 1)
```

[Return to Task](#Ch2PreClassQueT4)

#### PreClass Task 5

``` {r Ch2PreClassTask5, eval = FALSE}
group_by(pong_data2, BlockNumber, BackgroundColor)
```

[Return to Task](#Ch2PreClassQueT5)

#### PreClass Task 6

``` {r Ch2PreClassTask6, eval = FALSE}
pong_data2_group <- group_by(pong_data2, BackgroundColor, PaddleLength)
pong_data2_hits <- summarise(pong_data2_group, mean_hits = mean(HitOrMiss))
```

```{r lab2-preclass-sol-hidden-1, echo=FALSE}
pong_data2 <- filter(pong_data, TrialNumber >= 2) %>%
  mutate(TrialNumber = TrialNumber - 1) %>%
  mutate(Participant = Participant + 1)

# First we group the data accordingly, storing it in `pong_data2_group`
pong_data2_group <- group_by(pong_data2, BackgroundColor, PaddleLength)

# And then we summarise it, storing the answer in `total_hits`
pong_data2_hits <- summarise(pong_data2_group, mean_hits = mean(HitOrMiss))

# And then for fun we can filter just the red, small paddle hits
pong_data2_hits_red_small <- filter(pong_data2_hits, BackgroundColor == "red", PaddleLength == 50)
```

* You should find that the number of hits made with the small paddle (50) and the red color background is `r pong_data2_hits_red_small %>% pull(mean_hits)`

[Return to Task](#Ch2PreClassQueT6)

### InClass Actitivies

#### InClass Task 2

```{r Ch2InClassTask2-1, eval = FALSE}
library(tidyverse)
```

or

```{r Ch2InClassTask2-2, eval = FALSE}
library("tidyverse")
```

* Note, there is no difference between `library(tidyverse)` and `library("tidyverse")` both will work. 

[Return to Task](#Ch2InClassQueT2)

#### InClass Task 3

```{r Ch2InClassTask3, eval = FALSE}
responses <- read_csv("responses.csv")                  
qformats <- read_csv("qformats.csv")                 
scoring <- read_csv("scoring.csv")                  
pinfo <- read_csv("pinfo.csv")
```

* Note that there is a difference between `responses <- read_csv("responses.csv") ` and `responses <- read_csv(responses.csv) `. You will need quotes around the .csv filename as shown in the code chunk above (e.g. `responses <- read_csv("responses.csv")`), or the code won't work.

[Return to Task](#Ch2InClassQueT3)

#### InClass Task 7 

```{r Ch2InClassTask7, eval = FALSE}
aq_scores <- rscores %>% 
             group_by(Id) %>% # group by the ID number in column Id
             summarise(AQ = sum(Score)) # sum column Score to obtain AQ scores.
```

[Return to Task](#Ch2InClassQueT7)


#### InClass Task 8

```{r Ch2InClassTask8, eval = FALSE}
aq_scores2 <- responses %>% 
  pivot_longer(cols = Q1:Q10,
               names_to = "Question",
               values_to = "Response"),
  inner_join(qformats, "Question") %>%
  inner_join(scoring, c("QFormat", "Response")) %>%
             group_by(Id) %>% 
             summarise(AQ = sum(Score))
```

[Return to Task](#Ch2InClassQueT8)

### Assignment Solution

Solutions to the Assignment will appear here after Friday 30th October 2020. In the meantime please speak to Phil McAleer if you have any questions.

<span style="font-size: 22px; font-weight: bold; color: var(--purple);">Chapter Complete!</span>
