## Solutions to Questions

```{r lab3-solutions-data, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
menrot <-read_csv("./data/03-s01/inclass/MentalRotationBehavioralData.csv")
demog <- read_csv("./data/03-s01/inclass/demographics.csv")
``` 

Below you will find the solutions to the questions for the Activities for this chapter. Only look at them after giving the questions a good try and speaking to the tutor about any issues.

### InClass Activities

#### Task 1

`r hide("Solution Task 1")`
```{r, eval=FALSE}
library("tidyverse")

menrot <- read_csv("MentalRotationBehavioralData.csv")
demog <- read_csv("demographics.csv")
```
`r unhide()`

**Click the tab to see the solution**

#### Task 2

`r hide("Solution Task 2")`
```{r, eval=FALSE}
menrot_angle <- filter(menrot, CorrectResponse == "Correct") %>%
  inner_join(demog, "Participant") %>%
  group_by(Angle) %>% 
  summarise(mean_Resp = mean(Time))

ggplot(data = menrot_angle, aes(x = Angle, y = mean_Resp)) + 
  geom_point() +
  geom_line() +
  coord_cartesian(ylim = c(0, 3500), expand = TRUE)
```
`r unhide()`

**Click the tab to see the solution**

#### Task 3

`r hide("Solution Task 3")`
```{r, eval=FALSE}
menrot_angle_sex <- filter(menrot, CorrectResponse == "Correct") %>%
  inner_join(demog, "Participant") %>%
  group_by(Sex, Angle) %>% 
  summarise(mean_Resp = mean(Time))

ggplot(data = menrot_angle_sex, aes(x = Angle, y = mean_Resp)) + 
  geom_point() +
  geom_line() +
  coord_cartesian(ylim = c(0, 3500), expand = TRUE)
```
`r unhide()`

**Click the tab to see the solution**

#### Task 4

`r hide("Solution Task 4")`
```{r, eval=FALSE}
menrot_grouped <- filter(menrot, CorrectResponse == "Correct") %>%
  inner_join(demog, "Participant") %>%
  group_by(Sex, Angle) %>% 
  summarise(mean_Resp = mean(Time))

ggplot(data = menrot_grouped, aes(x = Angle, y = mean_Resp, group = Sex)) + 
  geom_point() +
  geom_line() +
  coord_cartesian(ylim = c(0, 3500), expand = TRUE)
```
`r unhide()`

**Click the tab to see the solution**

#### Task 5

`r hide("Solution Task 5")`
```{r, eval=FALSE}
menrot_grouped_color <- filter(menrot, CorrectResponse == "Correct") %>%
  inner_join(demog, "Participant") %>%
  group_by(Sex, Angle) %>% 
  summarise(mean_Resp = mean(Time))

ggplot(data = menrot_grouped_color, aes(x = Angle, y = mean_Resp, group = Sex)) + 
  geom_point(aes(color = Sex)) +
  geom_line() +
  coord_cartesian(ylim = c(0, 3500), expand = TRUE)
```
`r unhide()`

**Click the tab to see the solution**

#### Task 6

`r hide("Solution Task 6")`
```{r, eval=FALSE}
menrot_shape_size <- filter(menrot, CorrectResponse == "Correct") %>%
  inner_join(demog, "Participant") %>%
  group_by(Sex, Angle) %>% 
  summarise(mean_Resp = mean(Time))

ggplot(data = menrot_shape_size, aes(x = Angle, y = mean_Resp, group = Sex)) + 
  geom_point(aes(color = Sex, shape = Sex), size = 3) +
  geom_line() +
  coord_cartesian(ylim = c(0, 3500), expand = TRUE)
```
`r unhide()`

**Click the tab to see the solution**

#### Task 7

`r hide("Solution Task 7")`
```{r, eval=FALSE}
menrot_lab_theme <- filter(menrot, CorrectResponse == "Correct") %>%
  inner_join(demog, "Participant") %>%
  group_by(Sex, Angle) %>% 
  summarise(mean_Resp = mean(Time))

ggplot(data = menrot_lab_theme, aes(x = Angle, y = mean_Resp, group = Sex)) + 
  geom_point(aes(color = Sex, shape = Sex), size = 3) +
  geom_line() +
  coord_cartesian(ylim = c(0, 3500), expand = TRUE) +
  labs(x = "Angel of Rotation (degrees)", y = "Mean Reaction Time (ms)") +
  theme_bw()
```
`r unhide()`

**Click the tab to see the solution**

#### Task 8

`r hide("Solution Task 8")`
```{r, eval=FALSE}
manrot_facet_guide <- filter(menrot, CorrectResponse == "Correct") %>%
  inner_join(demog, "Participant") %>%
  group_by(Sex, Angle) %>% 
  summarise(mean_Resp = mean(Time))

ggplot(data = manrot_facet_guide, aes(x = Angle, y = mean_Resp, group = Sex)) + 
  geom_point(aes(color = Sex, shape = Sex), size = 3) +
  geom_line() +
  coord_cartesian(ylim = c(0, 3500), expand = TRUE) +
  labs(x = "Angel of Rotation (degrees)", y = "Mean Reaction Time (ms)") +
  facet_wrap(~Sex) +
  guides(group = FALSE, color = FALSE, shape = FALSE) +
  theme_bw()
```
`r unhide()`

**Click the tab to see the solution**
