## Solutions to Questions

Below you will find the solutions to the questions for the Activities for this chapter. Only look at them after giving the questions a good try and speaking to the tutor about any issues.

### InClass Activities

#### InClass Task 1

```{r, eval = FALSE}
ggplot(IQ_data, aes(IQ_data)) + 
  stat_function(fun = dnorm, args = list(mean = 100, sd = 10)) +
  labs(x = "IQ Score", y = "probability", title = "Distribution of IQ Scores") +
  theme_classic()
```


#### InClass Task 2

```{r, eval = FALSE}
ND_data <- tibble(ND_range = seq(-10,10,0.05))
ggplot(ND_data, aes(ND_range)) + 
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1)) +
  labs(x = "SD units", y = "probability", title = "The Normal Distribution") +
  theme_classic()
```

#### InClass Task 3

```{r, eval = FALSE}
# Key thing is to set lower.tail to FALSE to calculate the above area.
pnorm(190, 170, 7, lower.tail = FALSE)
```

#### InClass Task 4

```{r, eval = FALSE}
# the area is 185cm and above. 
# Key thing is to set lower.tail to FALSE to calculate the above area.
pnorm(185, 170, 7, lower.tail = FALSE)
```

#### InClass Task 5

```{r, eval = FALSE}
# Question 1
qnorm(0.95, 100, 15, lower.tail = TRUE)

# Question 2
  pnorm(105, 100, 15, lower.tail = FALSE) -
    pnorm(110, 100, 15, lower.tail = FALSE)
```


### InClass Binomial Activities

#### InClass Binomal CODE - Replace The Nulls

```{r, eval = FALSE}
blocks_5k <- replicate(n = 5000, 
                       sample(0:1, 
                              size = 60, 
                              replace = TRUE, 
                              c(2/3, 1/3)) %>% 
                         sum())
```

#### InClass Binomial Quickfire Questions Solutions

```{r, eval = FALSE}
# this time we switch it back to 1/3 prob because a face only has 1/3 prob of being shown
1. dbinom(10, 60, 1/3)

## NB: asks for more than 25, not 25 and more. Lower.tail is set to FALSE
2. pbinom(25, 60, 1/3, lower.tail = FALSE)

## lower.tail is true here becase you want lower than the cut-off.
qbinom(.05, 60, 1/3)
```

### Assignment Solution

**Part 1: Library**

There is a good chance that you will need the tidyverse library at some point in this exercise so load it in the code chunk below:

```{r library, eval = FALSE}
# hint: something to do with library() and tidyverse
library(tidyverse)
```

**Part 2: Basic probability and the binomial distribution**
 
You are conducting an auditory discrimination experiment in which participants have to listen to a series of sounds and determine whether the sound was human or not. On each trial participants hear one brief sound (100 ms) and must report whether the sound was from a human (coded as 1) or not (coded as 0). The sounds were either: a person, an animal, a vehicle, or a tone. 

#### Assignment Task 1

On a single trial, what is the probability that the sound will not be a person? Replace the `NULL` in the t1 code chunk with either mathematical notation or a single value. If entering a single value, give your answer to two decimal places.

```{r t1}
t1 <- 3/4
t1 <- 0.75

t1
```

#### Assignment Task 2

Over a sequence of 4 trials, with replacement, what is the probability of the following sequence of sounds: animal, animal, vehicle, tone? Replace the `NULL` in the t2 code chunk with either mathematical notation or a single value. If entering a single value, give your answer to three decimal places.

```{r t2}
t2 <- (1/4) * (1/4) * (1/4) * (1/4)
t2 <- .004

t2
```

#### Assignment Task 3

Over a sequence of four trials, without replacment, what is the probability of the following sequence of sounds: person, tone, animal, person? Replace the `NULL` in the t3 code chunk with either mathematical notation or a single value.

```{r t3}
t3 <- (1/4) * (1/3) * (1/2) * (0/1)
t3 <- 0

t3
```

#### Assignment Task 4

Replace the `NULL` below with code using the appropriate binomial distribution function to determine the probability of hearing 17 'tone' trials in a sequence of 100 trials. Assume the probability of a tone on any single trial is 1 in 4. Store the output in `t4`. 

```{r t4}
t4 <- dbinom(17, 100, 1/4)

t4
```

#### Assignment Task 5

Replace the `NULL` below with code using the appropriate binomial distribution function to determine what is the probability of hearing 30 'vehicle' trials or more in a sequence of 100 trials. Assume the probability of a vehicle trial on any one trial is 1 in 4. Store the output in `t5`. *Hint: do we want the upper or lower tails of the distribution*.

```{r t5}
t5 <- pbinom(29, 100, 1/4, lower.tail = FALSE)
t5 <- dbinom(30:100, 100, 1/4) %>% sum()

t5
```

#### Assignment Task 6

If a block of our experiment contained 100 trials, enter a line of code to run 10000 replications of one block, summing how many living sounds were heard in each replication. Code 1 for living sounds (person/animal) and 0 for non living sounds (vehicle/tone) and assume the probability of a living sound on any given trial is $p = .5$.

```{r t6}
t6 <- replicate(10000, sample(0:1, 100, TRUE, c(.5,.5)) %>% sum())

# Here we are only showing the first 10 values of 10000.
# Your numbers will vary from ours due to random sampling
# What would be assessed is the actual codeline exists and is accurate.
glimpse(t6)
```

**Part 2: The Normal Distribution**

Previously, in Lab 2, we looked at an ageing research project investigating differences in visual processing speed between younger (M = 22 years) and older adults (M = 71 years). One check in this experiment, prior to further analysis, is to make sure that older participants do not show signs of mild cognitive impairment (early symptoms of Alzheimer's disease). To do this we carry out a battery of cognitive tests to screen for such symptoms. One of the tests is the D2 test of attention which is a target cancellation task (i.e. participants cross out all letter d's with two dashes from a line of letters - see Lab 2 for an example). It is designed to test people's selective and sustained attention and visual scanning speed. The results of the test give a single score of *Concentration Performance* for each participant. The key piece of information for this analysis is that the distributions of D2 test scores are typically normally distributed (`M = 100`, `SD = 10`).  

#### Assignment Task 7

Replace the `NULL` below with code using the appropriate function to determine the probability that a given participant will have a D2 score of exactly 90? Store the output in `t7`

```{r t7}
t7 <- dnorm(90, 100, 10)

t7
```

#### Assignment Task 8

Replace the `NULL` below with code using the appropriate function to determine the probability that a given participant will have a D2 score of 120 or more? Store the output in `t8`

```{r t8}
t8 <- pnorm(120, 100, 10, lower.tail = FALSE)

t8
```

#### Assignment Task 9

Replace the `NULL` below with code using the appropriate function(s) to determine what is the difference in scores that cut off the top and bottom 5% of the distribution? Store the output in `t9`

```{r t9}
t9 <- qnorm(.95, 100, 10) - qnorm(.05, 100, 10)

t9
```

#### Assignment Task 10

Finally, if a participant says to you that they are worried because they have heard that their Concentration Performance was in the bottom 2% of scores on the distribution, what is the maximum D2 score that they can have? Replace the `NULL` below with a single value to two decimal places. Do not enter code. Store this in `t10`

```{r t10}
t10 <- 79.46

t10
```


<span style="font-size: 22px; font-weight: bold; color: var(--purple);">Chapter Complete!</span>