## Solutions to Questions

```{r lab5-solutions-data, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
RNGversion("3.5")
set.seed(1011)
dat <- read_csv("./data/05-s01/inclass/perm_data.csv") %>% 
  mutate(subj_id = row_number())
``` 

Below you will find the solutions to the questions for the Activities for this chapter. Only look at them after giving the questions a good try and speaking to the tutor about any issues.

### InClass Activities

#### Step 1

```{r data-load, message = FALSE, eval=FALSE}
library("tidyverse")

## Note: 
## Here's how we created the data for today's task:
## dat <- tibble(group = rep(c("A", "B"), each = 50),
##               Y = c(rnorm(50, 100, 15),
##                     rnorm(50, 110, 15)))
##
## write_csv(dat, "perm_data.csv")
## You could create a new dataset yourself and try it again.

dat <- read_csv("perm_data.csv") %>% 
  mutate(subj_id = row_number())
```

[Return to Task](#Ch5InClassQueT1)

#### Step 2

**Step 2.1.1 - the basic dat pipeline**

```{r basic-dat}
dat %>%
  group_by(group) %>%
  summarise(m = mean(Y))
```


**Step 2.1.2 - using `spread()` to separate the groups**
```{r spread-part}
dat %>%
  group_by(group) %>%
  summarise(m = mean(Y)) %>%
  spread(group, m)
```

**Step 2.1.3 - `mutate()` the column of mean differences**
```{r mutate-part}
dat %>%
  group_by(group) %>%
  summarise(m = mean(Y)) %>%
  spread(group, m) %>%
  mutate(diff = A - B)
```


**Step 2.1.4 - `pull()` out the difference**
```{r pull-part}
dat %>%
  group_by(group) %>%
  summarise(m = mean(Y)) %>%
  spread(group, m) %>%
  mutate(diff = A - B) %>%
  pull(diff)

```

**Step 2.2 - setting up the `calc_diff()` function**
```{r calc_diff_solution}
calc_diff <- function(x) {
  x %>%
    group_by(group) %>%
    summarise(m = mean(Y)) %>%
    spread(group, m) %>%
    mutate(diff = A - B) %>%
    pull(diff)
}
```

**Step 2.3 - Calculating d_orig using calc_diff()**
```{r d_orig_solution}
d_orig <- calc_diff(dat)
is.tibble(d_orig)
is.numeric(d_orig)
```

[Return to Task](#Ch5InClassQueT2)

#### Step 3

```{r permute_fn}
permute <- function(x) {
  x %>%
    mutate(group = sample(group))
}

permute(dat)
```

[Return to Task](#Ch5InClassQueT3)

#### Step 4

**Step 4.1 - the pipeline**
```{r permute_once}
dat %>% permute() %>% calc_diff()
```


**Step 4.2 - creating nhd**
```{r nhd}
nhd <- replicate(1000, dat %>% permute() %>% calc_diff())
```

**Step 4.3 - plotting nhd**
```{r histogram}
ggplot(tibble(x = nhd), aes(x)) + geom_histogram(binwidth = 1)
```

[Return to Task](#Ch5InClassQueT4)

#### Step 5

**Step 5.1 - The logical vector**

* This code establishes all the values in nhd that are equal to or greater than the value in d_orig
* It returns all these values as TRUE and all other values as FALSE
* `abs()` tells the code to ignore the sign of the value (i.e. assumes everything is positive)
```{r lvec}
lvec = abs(nhd) >= abs(d_orig)
```


**Step 5.2 - Sum up all the TRUE values**

* This gives the total number of values greater or equal to d_orig
```{r n-exceed}
n_exceeding_orig <- sum(lvec)
```

**Step 5.3 - Calculate the probability**

* The probability of obtaining d_orig or greater is calculated by the number of values equal to or greater than d_orig, divided by the full size of nhd (or in other words, its length)
```{r prob-p}
p <- n_exceeding_orig / length(nhd)
```

[Return to Task](#Ch5InClassQueT5)

<span style="font-size: 22px; font-weight: bold; color: var(--purple);">Chapter Complete!</span>
