## PreClass Activity 

We will now take each skill in turn. Be sure to try them all out. It looks a lot of reading but it is mainly just showing you the output of the functions so you can see you are doing it correctly. The key thing is to try them yourselves and don't be scared to change things to see what might happen if you do it slightly differently. We will also ask a couple of questions along the way to make sure you understand the skills.

### Skill 1: Generating Random Numbers

The `base::rnorm()` function generates values from a normal distribution and takes the following arguments:

  + `n`: the number of observations to generate
  + `mean`: the mean of the distribution (default 0)
  + `sd` : the standard deviation of the distribution (default 1)

To generate 10 or even 50 random numbers from a standard normal distribution (M = 0, SD = 1), you would use `rnorm(10)` or `rnorm(50)` respectively.  

* Type `rnorm(50)` into your console and see what happens. Use the below example for `rnorm(10)` to help you.  
* Try increasing `n` to 1000.  

```{r random10}
rnorm(10)
```
<br>
<span style="font-size: 22px; font-weight: bold; color: var(--green);">Quickfire Questions</span>  

If you enter `rnorm(50)` again you will get different numbers. Why? `r mcq(c("I have made a mistake",answer = "The numbers are random","R has made a mistake","Phil has made a mistake"))`

If you want to change the mean or sd, you would need to pass additional arguments to the function as shown below.  

```{r mean_example, eval = FALSE}
rnorm(n = 10, mean = 1, sd = .5)
```

* Try changing the mean and sd values a couple of times and see what happens. You get different numbers again that will be around the mean you set! Set a mean of 10, then a mean of 100, to test this.

Finally, for this Skill, you can concatenate (i.e. link) numbers together into a single vector using the `c()` function from base R.  For instance, say you wanted to create a vector with two sets of 50 random numbers from two separate samples: one set of 50 with a mean of 75 and the other with a mean of 90, you would use:

```{r random_numbers}
random_numbers <- c(rnorm(50, 75),
                    rnorm(50, 90))
```

<span style="font-size: 22px; font-weight: bold; color: var(--green);">Quickfire Questions</span>  

In the above example code, what is the standard deviation of the two samples you have created? `r mcq(c("50","75","90",answer = "1"))`

`r hide("Explain This - I don't get this answer!")`
```{block, type ="info"}
What is the **default** sd of the function?

Both populations would have an sd of 1, because that is the default, although you could easily change that.  Try it out!
```
`r unhide()` 
<br>
It is always good to check that your new vector has the right number of data points in it - i.e. the total of the two samples; a sanity check if you will. The new vector `random_numbers` should have `r length(random_numbers)` elements. You could verify this using the `length()` function:

```{r len_random_numbers}
length(random_numbers)
```
<br>
<span style="font-size: 18px; font-weight: bold; color: var(--blue);">Skill 1 out of 6 Complete!</span>

### Skill 2: Permuting Values

Another thing that is useful to be able to do is to generate **permutations** of values.  

`r hide("Portfolio Point - What are Permutations?")`
```{block, type ="info"}
A **permutation** is just a different ordering of the same values.  For example, the numbers 1, 2, 3 can be permuted into the following 6 sequences:

+ 1, 2, 3
+ 1, 3, 2
+ 2, 1, 3
+ 2, 3, 1
+ 3, 1, 2
+ 3, 2, 1

The more values you have, the more permutations of the order you have. The number of permutations can be calculated by, for example, `3*2*1`, where 3 is the number of values you have. Or through code: `factorial(3) = 6`. This assumes that each value is used once in the sequence and that each value never changes, i.e. 1234 cannot suddenly become 1235.
```
`r unhide()`

We can create random permutations of a vector using the `sample()` function. Let's use one of R's built in vectors: `letters`. 

* Type `letters` into the console, as below, and press RETURN/ENTER. You will see it contains all the lowercase letters of the English alphabet. Now, I bet you are wondering what `LETTERS` does, right?

```{r letters}
letters
```

We can combine `base::sample()` with `letters` to put the letters into a random order:

* Run the below line. Run it again. And again. What do you notice? And why is our output different from yours? (The answer is below)

```{r sample-letters}
sample(letters)
```

<span style="font-size: 22px; font-weight: bold; color: var(--green);">Quickfire Questions</span>  

If `month.name` contains the names of the twelve months of the year, how many possible permutations are there of `sample(month.name)`? `r mcq(c("1","12","144",answer = "479001600"))`

`r hide("Portfolio Point - Different samples with sample()")`
```{block, type ="info"}
Each time you run `sample(letters)` it will give you another random permutation of the sequence. That is what `sample()` does - creates a random permutation of the values you give it. Try repeating this command many times in the console. Because there are so many possible sequences, it is very unlikely that you will ever see the same sequence twice! 

An interesting thing about `sample()` is that `sample(c(1,2,3,4))` is the same as `sample(4)`. And to recap, there would be 24 different permutations based on `factorial(4)`, meaning that each time you type `sample(4)` you are getting one of those 24 different orders. So what would factorial(12) be?

Top Tip: Remember that you can scroll up through your command history in the console using the up arrow on your keyboard; this way, you don't ever have to retype a command you've already entered.  
```
`r unhide()`
<br>
<span style="font-size: 18px; font-weight: bold; color: var(--blue);">Skill 2 out of 6 Complete!</span>

### Skill 3: Creating Tibbles

Tables are important because most of the data we want to analyze comes in a table, i.e. tabular form. There are different ways to get tabular data into R for analysis.  One common way is to load existing data in from a data file (for example, using `readr::read_csv()` which you have seen before).  But other times, you might want to just type in data directly.  You can do this using the `tibble::tibble()` function. Being able to create a tibble is a useful data analysis skill because sometimes you will want to create some data on the fly just to try certain codes or functions. 

#### Entering Data into a Tibble

The `tibble()` function takes named arguments - this means that the name you give each argument within the tibble function, e.g. `Y = rnorm(10)` will be the name of the column that appears in the table, i.e. `Y`.  It's best to see how it works through an example.

```{r tibble-base}
tibble(Y = rnorm(10))
```

The above command creates a new table with one column named `Y`, and the values in that column are the result of a call to `rnorm(10)`: 10 randomly sampled values from a standard normal distribution (mean = 0, sd = 1) - See Skill 1.

If, however, we wanted to sample from two different populations for `Y`, we could combine two calls to `rnorm()` within the `c()` function. Again this was in Skill 1, here we are now just storing it in a tibble. See below:

```{r tibble1}
tibble(Y = c(rnorm(5, mean = -10), 
             rnorm(5, mean =  20)))
```

Now we have sampled a total of 10 observations - the first 5 come from a group with a mean of -10, and the second 5 come from a group with a mean of 20. Try changing the values in the above example to get an idea of how this works. Maybe even add a third group!

But, of course, it would be good to know which population each data point refers to and so we should add some group names. We can do this with some additional trickery using the `rep()` function.

#### Repeating Values to Save Typing

Before finalising our table, let's learn a little about the base R function, `rep()`.  This is most useful for automatically repeating values in order to save typing.  For instance, if we wanted 20 letter "A"s in a row, we would type:

```{r a_20}
rep("A", 20)
```

The first argument to `rep()` is the vector containing the information you want repeated, **A**, and the second argument, `times`, is the number of times to repeat it; in this case **20**.

If you wanted to add more information, e.g. if the first argument has more than one element, say "A" and "B", it will repeat the entire vector that number of times; A B, A B, A B, ... .  Note that we enclose "A" and "B" in the `c()` function so that it is seen as a single argument.

```{r A_and_B}
rep(c("A", "B"), 20)
```

But sometimes we want a specific number of As followed by a specific number of Bs; A A A B B B.  If the `times` argument has the same number of elements as the vector given for the first argument, it will repeat each element of the first vector as many times as given by the corresponding element in the `times` vector. In other words, for example, `times = c(2, 4)` for vector `c("A", "B")` will give you 2 As followed by 4 Bs.

```{r As_then_Bs}
rep(c("A", "B"), c(2, 4))
```

<span style="font-size: 22px; font-weight: bold; color: var(--green);">Quickfire Questions</span>

The best way to learn about this function is to play around with it in the console and see what happens. From the dropdown menus, the correct output of the following function would be:

1. `rep(c("A", "B", "C"),(2, 3, 1))`  - `r mcq(c("A A A B B C",answer = "A A B B B C","A A B B C C","A B C A B C"))`

2. `rep(1:5, 5:1)` - `r mcq(c("1 2 3 4 5 1 2 3 4 5 1 2 3 4 5","5 5 5 5 5 4 4 4 4 3 3 3 2 2 1",answer = "1 1 1 1 1 2 2 2 2 3 3 3 4 4 5","1 1 1 1 1 1 1 1 5 5 5 5 5 5 5"))`  
 
#### Bringing it Together in a Tibble

Now we know `rep()`, we can complete our table of simulated data by combining what we've learned about generating random numbers and repeating values. We want our table to look like this:

```{r tibble2, echo = FALSE}
tibble(group = rep(c("A", "B"), c(5, 5)),
       Y = c(rnorm(5, mean = -10), rnorm(5, mean =  20)))
```

You now know how to create this table. Have a look at the code below and make sure you understand it. We have one column called `group` where we create **A**s and **B**s through `rep()`, and one column called **Y**, our data, all in our `tibble()`:

```{r tibble3, results = "hide"}
tibble(group = rep(c("A", "B"), c(5, 5)),
       Y = c(rnorm(5, mean = -10), rnorm(5, mean =  20)))
```

Be sure to play around with the code chunk to get used to it. Try adding a third group or even a third column? Perhaps you want to give every participant a random age with a mean of 18, and an sd of 1; or even a participant number.

`r hide("Helpful Hint")`
```{block, type ="info"}
Try `row_number()` to create participant numbers.
```
`r unhide()` 
<br>
Don't forget, if you wanted to store your tibble, you would just assign it to a name, such as `my_data`:

```{r tibble3-new, results = "hide"}
my_data <- tibble(ID = row_number(1:10), 
                    group = rep(c("A", "B"), c(5, 5)),
                    Y = c(rnorm(5, mean = -10), rnorm(5, mean =  20)),
                    Age = rnorm(10, 18, 1))
```
<br>
<span style="font-size: 18px; font-weight: bold; color: var(--blue);">Skill 3 out of 6 Complete!</span>

### Skill 4: Computing Differences in Group Means

You have already learned how to calculate group means using `group_by()` and `summarise()`. For example, you might want to calculate sample means for a randomly generated dataset like so:

```{r tibble-long}
my_data <- tibble(group = rep(c("A", "B"), c(20, 20)),
                  Y = c(rnorm(20,  20, 5), rnorm(20, -20, 5)))

my_data_means <- my_data %>%
  group_by(group) %>%
  summarise(m = mean(Y))

my_data_means
```

Sometimes what we want though is to calculate **the differences between means** rather than just the means; so we'd like to subtract the second group mean `r my_data_means %>% pull(m) %>% pluck(2) %>% round(1)` from the first group mean of `r my_data_means %>% pull(m) %>% pluck(1) %>% round(1)`, to get a single value, the difference: `r (my_data_means %>% pull(m) %>% pluck(1) - my_data_means %>% pull(m) %>% pluck(2)) %>% round(1)`.

We can do this using the `dplyr::pull()` and `purrr::pluck()` functions.  `pull()` will extract a single column from a dataframe and turn it into a vector.  `pluck()` then allows you to pull out an element (i.e. a value or values) from within that vector.

```{r pull}
vec <- my_data_means %>%
  pull(m)

vec
```

We have now created `vec` which is a vector containing only the group means; the rest of the information in the table has been discarded.  Now that we have `vec`, we can calculate the mean difference as below, where `vec` is our vector of the two means and `[1]` and `[2]` refer to the two means:

```{r vec-diff}
vec[1] - vec[2]
```

But `pluck()` is also useful, and can be written as so: 

```{r pluck}
pluck(vec, 1) - pluck(vec, 2)
```

It can also be incorporated into a pipeline, as below, where we still `pull()` the means column, `m`, and then `pluck()` each value in turn and subtract them from each other.

```{r pull-pluck-pipe}
## whole thing in a pipeline
my_data_means %>% pull(m) %>% pluck(1) -
  my_data_means %>% pull(m) %>% pluck(2)
```

However, there is an alternative way to extract the difference between means which may make more intuitive sense.  You already know how to calculate a difference between values in the same row of a table using `dplyr::mutate()`, e.g. `mutate(new_column = column1 minus column2)`.  So if you can get the observations in `my_data_means` into the same row, different columns, you could then use `mutate()` to calculate the difference.  Previously you learned `gather()` to bring columns together. Well the opposite of gather is the `tidyr::spread()` function to split columns apart - as below.

```{r diff_spread}
my_data_means %>%
  spread(group, m)
```

The spread function (`?spread`) splits the data in column `m` by the information, i.e. labels, in column `group` and puts the data into separate columns.  A call to `spread()` followed by a `mutate()` can be used to calculate the difference in means - see below:

```{r spread_mutate}
my_data_means %>%
  spread(group, m) %>%
  mutate(diff = A - B) 
```

<span style="font-size: 22px; font-weight: bold; color: var(--green);">Quickfire Questions</span>  

What is the name of the column containing the differences between the means of A and B? `r mcq(c("means","group","m",answer = "diff"))`

Finally, if you then wanted to just get `diff` and throw away everything else in the table:

```{r spread_mutate_pull}
my_data_means %>%
  spread(group, m) %>%
  mutate(diff = A - B) %>%
  pull(diff)
```

`r hide("Portfolio Point - Reading pipes and verbalising tasks")`
```{block, type ="info"}
Keep in mind that a very useful technique for establishing what you want to do to a dataframe is to verbalise what you need, or to write it down in words, or to say it out loud. Take this last code chunk. What we wanted to do was to `spread()` the data in `m` into the groups A and B. Then we wanted to `mutate()` a new column that is the difference, `diff`, of A minus B. And finally we wanted to `pull()` out the value in `diff`. 

Often step 1 of writing code or understanding code is knowing what it is you want to do in the first place. After that, you just need the correct functions. Fortunately for us, `tidyverse` names its functions based on what they specifically do!
```
`r unhide()`
<br>
<span style="font-size: 18px; font-weight: bold; color: var(--blue);">Skill 4 out of 6 Complete!</span>

### Skill 5: Creating Your Own Functions

In Skills 1 to 4, we have looked at creating and sampling data, storing it in a tibble, and extracting information from that tibble. Now, say we wanted to do this over and over again. For instance, we might want to generate 100 random datasets just like the one in Skill 4. It would be a pain to have to type out the `tibble()` function 100 times or even to copy and paste it 100 times.  We'd likely make an error somewhere and it would be hard to read.  To help us, we can create a custom function that performs the action you want; in our case, creating a tibble of random data.  

Remember, a function is just a procedure that takes an input and gives you the same output each time - like a toaster! A function has the following format:

```{r function_example, eval = FALSE}
name_of_function <- function(arg1, arg2, arg3) {
  ## body of function goes between these curly brackets; i.e. what the function does for you.
  ## Note that the last value calculated will be returned if you call the function.
}
```

First, you define your own function name (e.g. `name_of_function`) and define the names of the arguments it will take (`arg1`, `arg2`, ...) - an argument is the information that you feed into your function, e.g. data. Finally, you state the calculations or actions of the function in the body of the function (the portion that appears between the curly braces). 

One of the most basic possible functions is one that takes no arguments and just prints a message. Here is an example:

```{r hello-world}
hello <- function() {
  print("Hello World!")
}
```

So this function is called `hello`. It can be run by typing `hello()` in your console and it will give the output of `Hello World!` every single time you run it; it has no other actions or information. Test this in the console now by typing:

```{r hello-call}
hello()
```

Awesome right? Ok, so not very exciting.  Let's make it better by adding an argument, `name`, and have it say Hello to `name`.

```{r hello-somebody}
hello <- function(name = "World!") {
  paste("Hello", name)
}
```

This new function is again called `hello()` and replaces the one you previously created. This time however you are supplying what is called a default argument, `name = "World!", but it still has the same action as the previous function of putting "Hello" and "World!" together. So if you run it you get "Hello World!". Try it yourself!

```{r hello-is-it-me-you-are-looking-for}
hello()
```

The difference this time however is that because you have added an argument to the input, you can change the information you give the argument and therefore change the output of the function. More flexible. More exciting.

<span style="font-size: 22px; font-weight: bold; color: var(--green);">Quickfire Questions</span>

Test your understanding by answering these questions:

* Typing `hello("Phil")` in the console with this new function will give: `r mcq(c("Hello Heather",answer = "Hello Phil","Hello Niamh","Hello Kevin"))`

* Typing the argument as `"is it me you are looking for"` will give: `r mcq(c(answer = "Hello is it me you are looking for","I just called to say Hello","You had me at Hello","Hello seems to be the hardest word"))`

* What argument would you type to get "Hello Dolly!" as the output: `r mcq(c("Dolly","Molly","Holly",answer = "Dolly!"))`

Most of the time however we want to create a function that computes a value or constructs a table.  For instance, let's create a function that returns randomly generated data from two samples, as we learned in the previous skills - see below. All we are doing is taking the tibble we created in Skill 4 and putting it in the body (between the curly brackets) of the function.

```{r gen-dat}
gen_data <- function() {
  tibble(group = rep(c("A", "B"), c(20, 20)),
                  Y = c(rnorm(20,  20, 5), rnorm(20, -20, 5)))
}
```

This function is called `gen_data()` and when we run it we get a randomly generated table of two groups, each with 20 people, one with M = 20, SD = 5, the other with M = -20, sd = 5. Try running this `gen_data()` function in the console a few times; remember that as the data is random, the numbers will be different each time you run it.  

But say we want to modify the function to allow us to get a table with smaller or larger numbers of observations per group.  We can add an argument `n` and modify the code as follows. Create this function and run it a couple of times through `gen_data()`. The way to think about this is that every place that `n` appears in the body of the function (between the curly brackets) it will have the value of whatever you gave it in the arguments, i.e. in this case, 20.

```{r gen-dat-20}
gen_data <- function(n = 20) {
  tibble(group = rep(c("A", "B"), c(n, n)),
                  Y = c(rnorm(n,  20, 5), rnorm(n, -20, 5)))
}
```

<span style="font-size: 22px; font-weight: bold; color: var(--green);">Quickfire Questions</span>

* How many total participants would there be if you ran `gen_data(2)`? `r mcq(c("2",answer = "4","20","40"))`

* What would you type to get 100 participants per group? `r mcq(c("gen_data(50)","gen_data(10)","gen_dota(100)",answer = "gen_data(100)"))`

**Challenge Question:**

Keeping in mind that functions can take numerous arguments, and that each group in your function have separate means, can you modify the function `gen_data` to allow the user to change the means for the two calls to `rnorm`? Have a try before revealing the solution below.

`r hide("Solution To Challenge Question")`
```{r gen-dat-min20, results = "hide"}
gen_data <- function(n = 20, m1 = 20, m2 = -20) {
  tibble(group = rep(c("A", "B"), c(n, n)),
                  Y = c(rnorm(n,  m1, 5), rnorm(n, m2, 5)))
}

# m1 is the mean of group A
# m2 is mean of group B

# The function would be called by: gen_data(20, 20, -20)
# Giving 20 participants in each group, 
# The first group having a mean of 20, 
# The second group having a mean of -20. 

# Likewise, a call of: gen_data(4, 10, 5)
# Would give two groups of 4, 
# The first having a mean of 10, 
# The second having a mean of 5.
```
`r unhide()`

`r hide("Portfolio Point - Two important facts about functions")`
```{block, type ="info"}
Here are two important things to understand about functions.

1. **Functions obey lexical scoping.**  What does this mean?  It's like what they say about Las Vegas: what happens in the function, stays in the function.  Any variables created inside of a function will be discarded after the function executes and will not be accessible to the outside calling process.  So if you have a line, say a variable `my_var <- 17` inside of a function, and try to print `my_var` from outside of the function, you will get an error: `object 'my_var' not found`. Although the function can 'read' variables from the environment that are not passed to it through an argument, it cannot change them.  So you can only write a function to return a value, not change a value.

2. **Functions return the last value that was computed.**  You can compute many things inside of a function but only the last thing that was computed will be returned as part of the calling process.  If you want to return `my_var`, which you computed earlier but not as the final computation, you can do so explicitly using `return(my_var)` at the end of the function (before the second curly bracket).
```
`r unhide()`
<br>
<span style="font-size: 18px; font-weight: bold; color: var(--blue);">Skill 5 out of 6 Complete!</span>

### Skill 6: Replicating Operations

The last skill you will need for the upcoming lab is knowing how to repeat an action (or expression) multiple times. You saw this in Lab 4 so we will only briefly recap here. Here, we use the  base function `replicate()`.  For instance, say you wanted to calculate the mean from `rnorm(100)` ten times, you could write it like this:

```{r reps, results = "hide"}
## bad way
rnorm(100) %>% mean()
rnorm(100) %>% mean()
rnorm(100) %>% mean()
rnorm(100) %>% mean()
rnorm(100) %>% mean()
rnorm(100) %>% mean()
rnorm(100) %>% mean()
rnorm(100) %>% mean()
rnorm(100) %>% mean()
```

But it's far easier and more readable to wrap the expression in the `replicate()` function where the first argument is the number of times you want to repeat the expression stated as the second argument, i.e. `replicate(times, expression)`. Below, we replicate the mean of 100 randomly generated numbers from the normal distribution, and we do this 10 times:

```{r rep10, results = "hide"}
replicate(10, rnorm(100) %>% mean())
```

Also you'll probably want to store the results in a variable, for example, `ten_samples`:

```{r 10-samples}
ten_samples <- replicate(10, rnorm(100) %>% mean())
ten_samples
```

Each element (value) of the vector within `ten_samples` is the result of a single call to `rnorm(100) %>% mean()`.

<span style="font-size: 22px; font-weight: bold; color: var(--green);">Quickfire Questions</span>  

Assuming that your `hello()` function from Skill 5 still exists, and it takes the argument `name = Goodbye`, what would happen in the console if you wrote, `replicate(1000, hello("Goodbye"))`? `r mcq(c("Hello World would appear a thousand times","hello Goodbye would appear a thousand times",answer = "Hello Goodbye would appear a thousand times"))` - Try it and see if it works!

`r hide("Solution To Quickfire Question")`
```{r sols-quick6, results = "hide"}
# the function would be:
hello <- function(name = "World!"){
  paste("Hello", name)
}

# and would be called by:
replicate(1000, hello("Goodbye"))
```
`r unhide()`
<br>

<span style="font-size: 18px; font-weight: bold; color: var(--blue);">Skill 6 out of 6 Complete!</span>

<span style="font-size: 22px; font-weight: bold; color: var(--blue);">Job Done - Activity Complete!</span>

To recap, we have shown you the following six skills:

  + Skill 1: Generating random numbers with `base::rnorm()` 
  + Skill 2: Permuting values with `base::sample()`  
  + Skill 3: Creating a "tibble" (a type of data table) using `tibble::tibble()`
  + Skill 4: Computing and extracting a difference in group means using `dplyr::pull()` and `purrr::pluck()`
  + Skill 5: Creating your own custom functions using `base::function()`
  + Skill 6: Repeating operations using `base::replicate()`

You will need these skills in the coming lab to help you perform a real **permutation test**. Through these skills and the permutation test you will learn about null hypothesis significance testing. If you have any questions please post them on the Moodle forum or the slack forum <a href = "https://rguppies.slack.com/" target = "_blank">rguppies.slack.com</a> under the channel **#level2_2018**. Finally, don't forget to add any useful information to your Portfolio before you leave it too long and forget. 

Great work today. That is all for now. See you in the lab!
