## Solutions to Questions

```{r lab12-solutions-setup, echo = FALSE, message=FALSE, warning=FALSE}
dmx <- tibble(sub_id = 1:9,
              i = rep(1:3, each = 3),
              j = rep(1:3, times = 3),
              typos = c(111, 102, 111, 
                    89, 127, 90,
                    97, 85, 88),
              sound = rep(c("cafe", "jazz", "silence"), each = 3)) %>%
  mutate(mu = mean(typos)) %>%
  group_by(i) %>%
  mutate(Ai = mean(typos) - mu) %>%
  ungroup() %>%
  mutate(err = typos - (mu + Ai))

dat_ss <- dmx %>% 
  summarise(total = sum(typos^2),
            ss_mu = sum(mu^2),
            ss_sound = sum(Ai^2),
            ss_err = sum(err^2))
```

Below you will find the solutions to the questions for the Activities for this chapter. Only look at them after giving the questions a good try and speaking to the tutor about any issues.

### InClass Activities

#### Task 1 Question 1

$SS_{total} - SS_{\mu}$ = `r as.integer(dat_ss[["total"]])` - `r as.integer(dat_ss[["ss_mu"]])` = `r as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])`

#### Task 1 Question 2

$SS_{sound} / SS_{corrected\ total}$ = `r as.integer(dat_ss[["ss_sound"]])` / `r as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])` = `r round(as.integer(dat_ss[["ss_sound"]]) / (as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])),2)`
    
#### Task 1 Question 3

$SS_{err} / SS_{corrected\ total}$ = `r as.integer(dat_ss[["ss_err"]])` / `r as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])` = `r round(as.integer(dat_ss[["ss_err"]]) / (as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])),2)`

#### Task 2 Question 1

k = number of conditions, so: 

df = k - 1 = 3 - 1 = 2

#### Task 2 Question 2

N = 9, and $dfA_{i}$ = 2, so

df = 9 - 2 - 1 = 6
    
#### Task 2 Question 3

```{r ms_a}
ms_a <- dat_ss[["ss_sound"]] / 2L  # 2 is the df for sound
    
ms_a
```

#### Task 2 Question 4

```{r ms_err}
ms_err <- dat_ss[["ss_err"]] / 6L  # 6 is the df for error

ms_err
```

#### Task 2 Question 5

$\mu_{cafe} = \mu_{jazz} = \mu_{silence}$

#### T3 Question 1

```{r fratio}
f_obs <- ms_a / ms_err
    
f_obs
```

#### T3 Question 2

The red line should be about F = 5.143, meaning that that is the minimum value of F, for those dfs, considered to be statistically significant at $\alpha = .05$

#### T3 Question 3

The approximate p-value for f_obs is p = .317.  

#### T3 Question 4

As $p$ > $\alpha$ we would retain the null hypothesis that there is no significant effect of group

#### T3 Q5 Hint

* look at inputs for the function - ?pf
* ignore ncp
* f_obs = q
* lower.tail? What is the probability of obtaining an F_obs higher than your value.

#### T3 Q5 Answer

```{r}
pf(f_obs, 2, 6, lower.tail = FALSE)
```

#### ezANOVA Solution

```{r ez, warning = FALSE}
dat <- select(dmx, sub_id, typos, sound)

ez::ezANOVA(dat,
            dv = typos,
            wid = sub_id,
            between = sound)
```