## Solutions to Questions

```{r lab12-solutions-setup, echo = FALSE, message=FALSE, warning=FALSE}
dmx <- tibble(sub_id = 1:9,
              i = rep(1:3, each = 3),
              j = rep(1:3, times = 3),
              typos = c(111, 102, 111, 
                    89, 127, 90,
                    97, 85, 88),
              sound = rep(c("cafe", "jazz", "silence"), each = 3)) %>%
  mutate(mu = mean(typos)) %>%
  group_by(i) %>%
  mutate(Ai = mean(typos) - mu) %>%
  ungroup() %>%
  mutate(err = typos - (mu + Ai))

dat_ss <- dmx %>% 
  summarise(total = sum(typos^2),
            ss_mu = sum(mu^2),
            ss_sound = sum(Ai^2),
            ss_err = sum(err^2))
```

Below you will find the solutions to the questions for the Activities for this chapter. Only look at them after giving the questions a good try and speaking to the tutor about any issues.

### InClass Activities

#### InClass Task 1 

**Question 1**

* $SS_{corrected\ total} = SS_{total} - SS_{\mu}$
* $SS_{total}$ = `r as.integer(dat_ss[["total"]])`
* $SS_{\mu}$ = `r as.integer(dat_ss[["ss_mu"]])`
* $SS_{corrected\ total}$ = `r as.integer(dat_ss[["total"]])` - `r as.integer(dat_ss[["ss_mu"]])` = `r as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])`

**Question 2**

* $SS_{sound}$ = `r as.integer(dat_ss[["ss_sound"]])`
* $SS_{corrected\ total}$ = `r as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])`
* $SS_{sound} / SS_{corrected\ total}$ = `r as.integer(dat_ss[["ss_sound"]])` / `r as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])` = `r round(as.integer(dat_ss[["ss_sound"]]) / (as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])),2)`
    
**Question 3**

* $SS_{err}$ = `r as.integer(dat_ss[["ss_err"]])`
* $SS_{corrected\ total}$ = `r as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])`
* $SS_{err} / SS_{corrected\ total}$ = `r as.integer(dat_ss[["ss_err"]])` / `r as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])` = `r round(as.integer(dat_ss[["ss_err"]]) / (as.integer(dat_ss[["total"]]) - as.integer(dat_ss[["ss_mu"]])),2)`

[Return to Task](#Ch12InClassQueT1)

#### InClass Task 2

**Question 1**

$\mu_{cafe} = \mu_{jazz} = \mu_{silence}$

**Question 2**

* k = number of conditions, so: 

df = k - 1 = 3 - 1 = 2

**Question 3**

* N = 9, and $dfA_{i}$ = 2, so

df = 9 - 2 - 1 = 6
    
**Question 4**

* The factor df = 2

```{r ms_a}
ms_a <- dat_ss[["ss_sound"]] / 2L
```

* Giving $MS_{A}$ = `r ms_a`

**Question 5**

* The residual df = 6

```{r ms_err}
ms_err <- dat_ss[["ss_err"]] / 6L
```

* Giving $MS_{S(A)}$ = `r ms_err`

[Return to Task](#Ch12InClassQueT2)

#### InClass Task 3

**Question 1**

```{r fratio}
f_obs <- ms_a / ms_err
```

* Giving a F-value of F = `r round(f_obs,3)` when rounded to three decimal places.

**Question 2**

The red line should be about F = 5.143, meaning that that is the minimum value of F, for those dfs, considered to be statistically significant at $\alpha = .05$

**Question 3**

Reading off the app, the approximate p-value for f_obs is p = .317.  

**Question 4**

As $p$ > $\alpha$ we would retain the null hypothesis that there is no significant effect of group in this study

**Question 5**

This could alternatively calculated using coding as such:

```{r ch12Task3Q5sol, eval=FALSE}
pf(f_obs, 2, 6, lower.tail = FALSE)
```

Which would give a p-value of p = `r pf(f_obs, 2, 6, lower.tail = FALSE) %>% round(3)` and again could be written up as F(2, 6) = `r f_obs %>% round(3)`, p = `r pf(f_obs, 2, 6, lower.tail = FALSE) %>% round(3)`, suggesting no significant effect of ambiance or location on concentration as measured through the number of typos.

[Return to Task](#Ch12InClassQueT3)

#### InClass Task 4

```{r ez, warning = FALSE}
dat <- select(dmx, sub_id, typos, sound)

results <- ez::ezANOVA(dat,
                       dv = typos,
                       wid = sub_id,
                       between = sound)
```

You didn't need to store the output but we have done so out of habit, as often we want to do something with the results.  And remember we said that you get two tables in this output.  One of the tables is Levene's Test for Homogeneity of Variance and it looks like:

```{r ez-output1, echo=FALSE, results='asis'}
knitr::kable(results$`Levene's Test for Homogeneity of Variance`,
             digits = 3,
             caption = "Levene's test for Homogeneity of Variance")%>%
  kable_styling() %>%
  scroll_box(width = "100%", box_css = "border: 0px;")
```
<br>
The other table is the main ANOVA output. Make sure you can locate both individual tables and you know which is which.

```{r ez-output2, warning = FALSE, echo=FALSE}
knitr::kable(results$ANOVA,
             digits = 3,
             caption = "ANOVA output")%>%
  kable_styling() %>%
  scroll_box(width = "100%", box_css = "border: 0px;")
```
<br>
From your lectures, remember that Levene's test relates to an assumption of equal variance across conditions in a between-subjects design. A non-significant finding in Levene's suggests equal variance across conditions as required by the assumptions of the ANOVA. The ANOVA is the actual output of the test and will match closely with the values you have found in Task 3. 

[Return to Task](#Ch12InClassQueT4)

<span style="font-size: 22px; font-weight: bold; color: var(--purple);">Chapter Complete!</span>
