## Solutions to Questions

```{r lab15-solutions-setup, echo = FALSE, message=FALSE, warning=FALSE}

pinfo <- read_csv("data/15-s02/inclass/participant_info.csv")
wellbeing <- read_csv("data/15-s02/inclass/wellbeing.csv")
screen <- read_csv("data/15-s02/inclass/screen_time.csv")
```

Below you will find the solutions to the questions for the PreClass and InClass activities for this chapter. Only look at them after giving the questions a good try and speaking to the tutor about any issues.

### PreClass Activities

#### Loading the Data

```{r ch15-sol-load, eval = FALSE}
library("tidyverse")

pinfo <- read_csv("participant_info.csv")
wellbeing <- read_csv("wellbeing.csv")
screen <- read_csv("screen_time.csv")
```

[Return to Task](#Ch15PreClassQueT1)

#### Compute the well-being score for each participant

* Both of these solutions would produce the same output.

**Version 1**

* bit quicker in terms of coding and reduces chance of error by perhaps forgetting to include a specific column 

```{r ch15-sol-wemwbs}
wemwbs <- wellbeing %>%
  gather("var", "score", -Serial) %>%
  group_by(Serial) %>%
  summarise(tot_wellbeing = sum(score))
```

**Version 2**

* this one is a bit slower

```{r ch15-sol-wemwbs2}
wemwbs <- wellbeing %>%
  mutate(tot_wellbeing = WBOptimf +	WBUseful + 
           WBRelax + WBIntp +	WBEnergy + WBDealpr + 
           WBThkclr + WBGoodme + WBClsep + WBConfid + 
           WBMkmind + WBLoved + WBIntthg + WBCheer) %>%
  select(Serial, tot_wellbeing) %>%
  arrange(Serial)
```

[Return to Task](#Ch15PreClassQueT2)

#### Visualising Screen time on all technologies

```{r ch15-sol-st-plot-sol, fig.cap = 'Count of the hours of usage of different types of social media at Weekdays and Weekends'}
## screen time
screen_long <- screen %>%
  gather("var", "hours", -Serial) %>%
  separate(var, c("variable", "day"), "_")

screen2 <- screen_long %>%
  mutate(variable = recode(variable,
			   "Watch" = "Watching TV",
			   "Comp" = "Playing Video Games",
			   "Comph" = "Using Computers",
			   "Smart" = "Using Smartphone"),
	 day = recode(day,
		      "wk" = "Weekday",
		      "we" = "Weekend"))

ggplot(screen2, aes(hours)) +
  geom_bar() +
  facet_grid(day ~ variable)
```

[Return to Task](#Ch15PreClassQueT3)

#### Visualising the Screen time and Well-being relationship

```{r ch15-sol-combined, fig.cap = 'Scatterplot showing the relationship between screen time and mean well-being across four technologies for Weekdays and Weekends'}
dat_means <- inner_join(wemwbs, screen2, "Serial") %>%
  group_by(variable, day, hours) %>%
  summarise(mean_wellbeing = mean(tot_wellbeing))

ggplot(dat_means, aes(hours, mean_wellbeing, linetype = day)) +
  geom_line() +
  geom_point() +
  facet_wrap(~variable, nrow = 2)
```

[Return to Task](#Ch15PreClassQueT4)

### InClass Activities

#### Smartphone and well-being for boys and girls

**Solution Steps 1 to 2**

```{r ch15-sol-calculate_mean}
smarttot <- screen2 %>%
  filter(variable == "Using Smartphone") %>%
  group_by(Serial) %>%
  summarise(tothours = mean(hours))
```

**Solution Step 3**

```{r ch15-sol-combine2}
smart_wb <- smarttot %>%
  filter(tothours > 1) %>%
  inner_join(wemwbs, "Serial") %>%
  inner_join(pinfo, "Serial") 
```

[Return to Task](#Ch15InClassQueT1)

#### Visualise and Interpreting the relationship of smartphone use and wellbeing by sex 

**The Figure**

```{r ch15-use-wb-bysex, fig.cap='Scatterplot and slopes for relationships between total hours and mean wellbeing score for boys (cyan) and girls (red)'}
smart_wb_gen <- smart_wb %>%
  group_by(tothours, sex) %>%
  summarise(mean_wellbeing = mean(tot_wellbeing))

ggplot(smart_wb_gen, aes(tothours, mean_wellbeing, color = factor(sex))) +
  geom_point() +
  geom_smooth(method = "lm")
```

**A brief Interpretation**

Girls show lower overall well-being compared to boys.  In addition, the slope for girls appears more negative than that for boys; the one for boys appears relatively flat.  This suggests that the negative association between well-being and smartphone use is stronger for girls.

[Return to Task](#Ch15InClassQueT2)

#### Estimating model parameters

* This is the chunk we gave in the materials.

```{r ch15-sol-centering}
smart_wb <- smart_wb %>%
  mutate(tothours_c = tothours - mean(tothours),
         sex_c = ifelse(sex == 1, .5, -.5)) %>%
  select(-tothours, -sex)
```

* and the model would be specified as:

```{r ch15-sol-mod}
mod <- lm(tot_wellbeing ~ tothours_c * sex_c, smart_wb)
```

* or alternatively

```{r ch15-sol-mod-option}
mod <- lm(tot_wellbeing ~ tothours_c + sex_c + tothours_c:sex_c, smart_wb)
```

* and the output called by:

```{r ch15-sol-mod-output}
summary(mod)
```

[Return to Task](#Ch15InClassQueT3)

#### Final Interpretations

```{r ch15-sol-hidden, include = FALSE}
male_intercept <- coef(mod)["(Intercept)"] + coef(mod)["sex_c"] * .5
male_slope <- coef(mod)["tothours_c"] + coef(mod)["tothours_c:sex_c"] * .5
female_intercept <- coef(mod)["(Intercept)"] + coef(mod)["sex_c"] * -.5
female_slope <- coef(mod)["tothours_c"] + coef(mod)["tothours_c:sex_c"] * -.5
```

* The interaction between smartphone use and gender is shown by the variable tothours_c:sex_c, and this interaction was significant at the $\alpha = .05$ level, meaning that there is an significant interaction between sex and hours of smartphone usage on wellbeing

* To two decimal places, the intercept for male participants is: `r round2(male_intercept, 2)`
* To two decimal places, the slope for male participants is: `r round2(male_slope, 2)`
* To two decimal places, the intercept for female participants is: `r round2(female_intercept, 2)`
* To two decimal places, the slope for female participants is: `r round2(female_slope, 2)`

As such, given the model of `Y = intercept + (slope * X)`  where Y is wellbeing and X is total hours on smartphone, what would be the predicted wellbeing score for a male and a female who use their smartphones for 8 hours. Give your answer to two decimal places:

* Male: `r round2(male_intercept, 2) + round2(male_slope, 2) * 8`
* Female: `r round2(female_intercept, 2) + round2(female_slope, 2) * 8`

And finally, the most reasonable interpretation of these results is that smartphone use was more negatively associated with wellbeing for girls than for boys.

[Return to Task](#Ch15InClassQueT4)

<span style="font-size: 22px; font-weight: bold; color: var(--purple);">Chapter Complete!</span>
